---
title: "01 Exploration"
author: '35762384'
date: "20/01/2021"
output: 
  html_document: 
    keep_md: yes
    code_folding: show
---

```{r, message=FALSE}
library(sf)
library(readxl)
library(dplyr)
library(plyr)
library(ggplot2)
library(afrihealthsites)
library(ggpubr)
library(afriadmin)
library(tmap)
library(cowplot)
library(forcats)
library(colorspace)
library(readr)

# Install Malawi MFL
malawi_MFL = read_excel("~/malawi-health-facilities-1/MHFR_Facilities 1.xlsx")

# Convert to sf

## omit NA's
new_malawi_MFL = na.omit(malawi_MFL)

## check for NA 
any(is.na(new_malawi_MFL))

## transform geometry columns into numeric 
sapply(new_malawi_MFL, class)
new_malawi_MFL = transform(new_malawi_MFL, LATITUDE = as.numeric(LATITUDE), 
                                           LONGITUDE = as.numeric(LONGITUDE))

## convert to sf object
malawi_facilities_MFL = st_as_sf(new_malawi_MFL, coords = c("LONGITUDE", "LATITUDE"), dim = "XY")

malawi_facilities_MFL = st_set_crs(malawi_facilities_MFL, 4326) ## set CRS

head(malawi_facilities_MFL)
```



Overview:

1. Malawi MFL
- Facility locations managed by the Ministry of Health and Population Malawi. WHO guidance states that it should be updated at least every 2 years. Last update is unknown. 
- Locations in lat-long, CRS set to WGS84
- 1546 facilities reported, data includes name, ownership, type of facility and functional status.
- 1426 facilities after removing NAs
- 9 types of facilities, including unclassified. Number of each type is shown. Top 3 are clinics (585), health centers (520) and Dispensaries (179). All are primary level. What is private? 
  
```{r, class.source= "fold-hide"}
# Re-order the facility types
malawi_MFL$TYPE = as.factor(malawi_MFL$TYPE)

malawi_MFL$TYPE = factor(malawi_MFL$TYPE, levels = c("Central Hospital", "District Hospital", "Hospital", "Health Centre", "Clinic", "Health Post", "Dispensary", "Private", "Unclassified"))

# Number of each type + ownership
facility_types_MFL = as.data.frame(table(malawi_MFL$TYPE, malawi_MFL$OWNERSHIP))
head(facility_types_MFL)
levels(facility_types_MFL$Var2)

# Only showing government, private and CHAM ownership
facility_types_MFL$new_Var2 = revalue(facility_types_MFL$Var2, c("Aquaid Lifeline"="Other", "Non-Government"="Other", "Parastatal"="Other", "Mission/Faith-based (other than CHAM)"="Other"))
facility_types_MFL$new_Var2 = factor(facility_types_MFL$new_Var2, levels = c("Government", "Private", "Christian Health Association of Malawi (CHAM)", "Other"))


## bar plot of no. of facility types 
plot_facility_types_MFL = ggplot(facility_types_MFL, aes(x=Var1, y=Freq, fill=new_Var2)) + geom_bar(position = "stack", stat = "identity")

plot_facility_types_MFL = plot_facility_types_MFL + labs(x = "Facility types", y = "Frequency", fill="Ownership") + scale_fill_brewer(palette = "Set2") + coord_flip() + theme_minimal() + ggtitle("MFL") + theme(axis.title.y = element_blank(), legend.title = element_text(size = 8, face = "bold"), legend.text = element_text(size = 8), plot.title = element_text(face = "bold", hjust = 0), legend.position = "bottom")
plot_facility_types_MFL

par(mar=c(11,4,4,4))
```
  
  
- 8 types of ownership, top 3 are government (695), Private (495) and CHAM. It seems that private not for profit is separated from private for profit in this data set. 

```{r, class.source = 'fold-hide'}
# Re-order ownership
malawi_MFL$OWNERSHIP = as.factor(malawi_MFL$OWNERSHIP)

malawi_MFL$OWNERSHIP = factor(malawi_MFL$OWNERSHIP, levels = c("Government", "Private", "Christian Health Association of Malawi (CHAM)", "Non-Government", "Mission/Faith-based (other than CHAM)", "Other", "Parastatal", "Aquaid Lifeline"))

# Number of each type of ownership
ownership_MFL = as.data.frame(table(malawi_MFL$OWNERSHIP))

## bar plot of ownership
plot_ownership_MFL = ggplot(ownership_MFL, aes(x=forcats::fct_relabel(Var1,stringr::str_wrap,width = 16), y=Freq)) + geom_bar(stat = "identity", fill="slategray")

plot_ownership_MFL = plot_ownership_MFL + labs(x="Ownership", y = "Frequency") + coord_flip() + theme_minimal() + ggtitle("MFL") + theme(axis.title.y = element_blank(), plot.title = element_text(face = "bold"))
plot_ownership_MFL

```



2. WHO-KWTRP
- Focuses on facilities run by government, faith-based organisations, NGO's and local authorities. Covers 50 countries in sub-Saharan Africa. Sources of information include health sector reports, websites run by national or international organisations and personal communications
- If MFL was available it was used. More than one datasource was often used to compile facility list
- Private facilities are excluded, duplicates removed, name errors corrected and name variations were matched. Missing info was added with the use of other datasources. 
- Now hosted by the WHO Global Malaria Programme, last update February 2019
   
- Malawi datasources includes MFL, https://data.humdata.org/dataset/malawi-health and http://www.cham.org.mw/uploads/7/3/0/8/73088105/cham_health_facilities_-_1_june_2016.pdf
- At time of publishing, 639 facilities with 9 missing coordinates, not been updated since
- Data includes facility name, type, ownership, source of location and reclassified facility types

```{r, class.source = 'fold-hide'}

# Malawi WHO data.frame
malawi_WHO <- afrihealthsites("malawi", datasource='who', plot=FALSE, returnclass='dataframe')

head(malawi_WHO)

```


- Original facility types: 8 types, top 3 are health centers (457), health post/dispensary (87) and mission hospitals (27). Large difference between types. 
   
```{r, class.source = 'fold-hide'}
# Re order facility types and ownership 
malawi_WHO$`Facility type` = as.factor(malawi_WHO$`Facility type`)
malawi_WHO$`Facility type` = factor(malawi_WHO$`Facility type`, levels = c("Central Hospital", "District Hospital", "Mission Hospital", "Rural Hospital", "Community Hospital", "Health Centre", "Clinic", "Health Post/Dispensary"))

malawi_WHO$Ownership = as.factor(malawi_WHO$Ownership)
malawi_WHO$Ownership = factor(malawi_WHO$Ownership, levels = c("MoH", "FBO", "Local authority", "NGO"))

# No. of original facility types + ownership
facility_types_WHO = as.data.frame(table(malawi_WHO$`Facility type`, malawi_WHO$Ownership))
head(facility_types_WHO)

## bar plot of original facility types 
plot_facility_types_WHO = ggplot(facility_types_WHO, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(position = "stack", stat = "identity")
plot_facility_types_WHO = plot_facility_types_WHO + labs(x = "Facility types", y = "Frequency", fill="Ownership") + scale_fill_brewer(palette = "Set2") + coord_flip() + theme_minimal() + ggtitle("WHO") + theme(axis.title.y = element_blank(), legend.title = element_text(size = 8, face = "bold"), legend.text = element_text(size = 8), plot.title = element_text(face = "bold", hjust = 0), legend.position = "bottom") + expand_limits(y=c(0,600))
plot_facility_types_WHO

```
 
   
- Reclassified facility types: Hospitals are aggregated except Community Hospital (became Community Health Unit), the rest are the same. 
   
```{r, class.source = 'fold-hide'}
# Re order
malawi_WHO$facility_type_9 = as.factor(malawi_WHO$facility_type_9)
malawi_WHO$facility_type_9 = factor(malawi_WHO$facility_type_9, levels = c("Hospital", "Health Centre", "Health Clinic", "Health Post", "Community Health Unit"))

# No. of reclassified facility types
RC_facility_types_WHO = as.data.frame(table(malawi_WHO$facility_type_9))
RC_facility_types_WHO

## bar plot of reclassified facility types 
plot_RC_facility_types_WHO = ggplot(RC_facility_types_WHO, aes(x=Var1, y=Freq)) + geom_bar(stat = "identity", fill="slategray")
plot_RC_facility_types_WHO = plot_RC_facility_types_WHO + labs(x = "Reclassified facility types", y = "Frequency") + coord_flip() + theme_minimal() + theme(axis.title.y = element_blank(), plot.title = element_text(face = "bold")) + ggtitle("WHO") + expand_limits(y=c(0,600))
plot_RC_facility_types_WHO
```
   
   
- Ownership: 4 types, majority is government owned (467) with faith based organisations second. 
   
```{r, class.source = 'fold-hide'}
# Types of ownership
ownership_WHO = as.data.frame(table(malawi_WHO$Ownership))
ownership_WHO

## bar plot of ownership
plot_ownership_WHO = ggplot(ownership_WHO, aes(x=Var1, y=Freq)) + geom_bar(stat = "identity", fill="slategray")
plot_ownership_WHO = plot_ownership_WHO + labs(x="Ownership", y = "Frequency") + coord_flip() + theme_minimal() + ggtitle("WHO") + theme(axis.title.y = element_blank(), plot.title = element_text(face = "bold")) + expand_limits(y=c(0,600))
plot_ownership_WHO
```



Both data sources contain no information on services available, capacity or equipment. MFL does state whether facility is functional.

Classification of MFL facilities aligns more with the structure of the health care system in Malawi (community, primary, secondary, tertiary), it differentiates central hospitals from district and other hospitals. WHO has additional rural and mission hospitals, where do they fit in?

https://www.health.gov.mw/index.php/2016-01-06-19-58-23/national-aids states that at community level, health posts, dispensaries and maternity clinics offer services. Primary includes health centers and community hospitals, secondary consists of district and some CHAM hospitals, tertiary includes central hospitals. 



Analysis:

- 1546 compared to 648 facilities, difference of 898

```{r, echo=FALSE}
# facility types
plot_facilities = align_plots(plot_facility_types_MFL, plot_facility_types_WHO, align = "v")
ggdraw(plot_facilities[[1]])
ggdraw(plot_facilities[[2]])
```


- Ownership

```{r, echo=FALSE}
# ownership
plot_ownership = align_plots(plot_ownership_MFL, plot_ownership_WHO, align = "v")
ggdraw(plot_ownership[[1]])
ggdraw(plot_ownership[[2]])
```


- Maps

```{r, echo=FALSE, fig.show='hide'}

# choose admin level
malawi_admin2 = afriadmin("malawi",level=2, plot='sf')
malawi_admin1 = afriadmin("malawi", level = 1, plot = "sf")

# static WHO facility location map
map_static_WHO = afrihealthsites("malawi", datasource='who', plot='sf')

```

```{r, echo=FALSE}
### changing character class to factor for the categorical variable column 

### MFL

malawi_facilities_MFL$TYPE = as.factor(malawi_facilities_MFL$TYPE)

# re-ordering the levels 
malawi_facilities_MFL$TYPE = factor(malawi_facilities_MFL$TYPE, levels = c("Central Hospital", "District Hospital", "Hospital", "Health Centre", "Clinic", "Health Post", "Dispensary", "Private", "Unclassified")) 

### WHO
map_static_WHO$`Facility type` = as.factor(map_static_WHO$`Facility type`)

# re-ordering the levels
map_static_WHO$`Facility type` = factor(map_static_WHO$`Facility type`, levels = c("Central Hospital", "District Hospital", "Mission Hospital", "Rural Hospital", "Community Hospital", "Health Centre", "Clinic", "Health Post/Dispensary"))

# tmap

tmap_mode("view")

## admin map
tmap_admin2 = tm_shape(st_geometry(malawi_admin2)) + tm_borders()

## MFL facility locations
tmap_facilities_MFL = tmap_admin2 + tm_shape(malawi_facilities_MFL) + tm_dots(col = "TYPE", palette = "-viridis") + tm_layout(frame = FALSE, asp = 2, title = "MFL", title.position = c("left", "top"))
tmap_facilities_MFL

## WHO facility locations
tmap_facilities_WHO = tmap_admin2 + tm_shape(map_static_WHO) + tm_dots(col = "Facility type", palette = "-viridis") + tm_layout(frame = FALSE, asp = 2, title = "WHO", title.position = c("left", "top"))
tmap_facilities_WHO


## final 
tmap_arrange(tmap_facilities_MFL, tmap_facilities_WHO, ncol = 2) # side by side

tmap_facilities_MFL_WHO = tmap_facilities_WHO + tm_shape(malawi_facilities_MFL) + tm_dots(col = "TYPE", palette = "-YlOrRd", alpha = 0.7) + tm_layout(frame = FALSE, asp = 2, title = "MFL & WHO", title.position = c("left", "top")) 
tmap_facilities_MFL_WHO # combined

```

  Qs to address?:
  
  1. How many facilities are in the same location across the MFL and WHO datasets?
  
  - 2 points intersect directly and so are same up to 5 decimal places (WHO data has all coordinates to 5 decimal places). Check fewer decimal places to see if more intersect? 
  
  2. Do same facilities share same names/other attributes?
  
  - Ownership is the same for both facilities but Euthini is registered as a rural hospital in the MFL data while in the WHO data, it is a health centre. The name reflects this. 
  

```{r}
# Qs 1 - how many intersect?

## convert malawi_WHO to sf object

class(malawi_WHO)
any(is.na(malawi_WHO))

new_malawi_WHO = na.omit(malawi_WHO) ## omit NA

sf_malawi_WHO = st_as_sf(new_malawi_WHO, coords = c("Long", "Lat"), dim = "XY")
sf_malawi_WHO = st_set_crs(sf_malawi_WHO, 4326)

## st_intersection
intersect_WHO_MFL = st_intersection(x=sf_malawi_WHO, y=malawi_facilities_MFL)
intersect_WHO_MFL ## only 2 intersect directly, so are same up to 5 decimal places?
```



  3. The ones that aren't, are they within 50m of another facility?
  
```{r}
# Try merge_points()
merge_points("malawi", datasources=c('who', 'healthsites'), dist_same_m=50)

```



  4. How many facilities per admin1 and admin2 regions?
  
  - Top 3 districts for MFL are: Lilongwe (163 facilities), Blantyre (154) and Mzimba (95)
  - Top 3 districts for WHO are: Mzimba (56 facilities), Lilongwe (52) and Mangochi (45)

```{r, message=FALSE, warning=FALSE}

# Facilities per region 
# MFL, admin1
facility_admin1_MFL = st_intersects(malawi_admin1, malawi_facilities_MFL, sparse = TRUE)
malawi_admin1$facility = lengths(facility_admin1_MFL)
head(malawi_admin1) # gave no. of facilities per region

# Intersect opposite to get admin1 region for each facility 
admin1_facility_MFL = unlist(st_intersects(malawi_facilities_MFL, malawi_admin1, sparse = TRUE))

# error when adding admin1_facility_MFL data to malawi_facilities_MFL, 
# no. of rows are less in admin1_, why?


# trying to figure out which points don't intersect 

intersect_admin1_MFL = st_intersection(malawi_facilities_MFL, malawi_admin1) # returns data frame of matches

# working out which facilities did not intersect
intersect_admin1_MFL = as.data.frame(intersect_admin1_MFL) # convert to data frame to use in anti_join

no_intersect_admin1_MFL = anti_join(malawi_facilities_MFL, intersect_admin1_MFL) # shows which did not intersect 

# turn back into sf 
no_intersect_admin1_MFL = st_as_sf(no_intersect_admin1_MFL) 
intersect_admin1_MFL = st_as_sf(intersect_admin1_MFL) 

# !!! something is wrong, facilities with same district show different 
# value for region. Was data for district inputted incorrectly or are 
# coordinates wrong?


## Trying to find the source of the error 

## adding admin1_facility_MFL was the issue (the results from st_intersect),
## st_intersect and st_intersection are not computed in the same order
## just use intersect_admin1_MFL for the plots, don't need to combine with facilities that did not
## intersect


# Map with admin1 layer 
tmap_admin1_MFL = tmap_facilities_MFL + tm_shape(st_geometry(malawi_admin1)) + tm_borders()
tmap_admin1_MFL

## Map - number of facilities per region
tmap_admin1_MFL2 = tm_shape(st_geometry(malawi_admin1)) + tm_borders() + tm_shape(malawi_admin1) + tm_fill("facility", style = "cat", palette = sequential_hcl(4, palette = "YlGnBu", rev = TRUE), title = "MFL facilities")

tmap_admin1_MFL2

# Plot of facilities in each admin1 region 

## transform columns
intersect_admin1_MFL$shapeName = as.factor(intersect_admin1_MFL$shapeName)
intersect_admin1_MFL$TYPE = as.factor(intersect_admin1_MFL$TYPE)

## freq of facility types by region
df_facility_admin1_MFL = as.data.frame(table(intersect_admin1_MFL$shapeName, 
                                             intersect_admin1_MFL$TYPE))

## subsetting by region for plots 

# region 1 = Central 
admin1_region1 = filter(df_facility_admin1_MFL, Var1 == "Central Region")

admin1_region1_plot = ggplot(admin1_region1, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "MFL Central Region (total facilities=1365)") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0, 250))

# region 2 = Southern
admin1_region2 = filter(df_facility_admin1_MFL, Var1 == "Southern Region")

admin1_region2_plot = ggplot(admin1_region2, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "MFL Southern Region (total facilities=1365)") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0,250))

# region 3 = Northern
admin1_region3 = filter(df_facility_admin1_MFL, Var1 == "Northern Region")

admin1_region3_plot = ggplot(admin1_region3, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "MFL Northern Region (total facilities=1365)") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0, 150))


# MFL, admin2 
facility_admin2_MFL = st_intersects(malawi_admin2, malawi_facilities_MFL, sparse = TRUE)
malawi_admin2$facility = lengths(facility_admin2_MFL)

sum(malawi_admin2$facility) # also has 1365 intersections

# Map - number of facilities per district
tmap_admin2_MFL = tm_shape(st_geometry(malawi_admin2)) + tm_borders() + tm_shape(malawi_admin2) + tm_fill("facility", palette = sequential_hcl(6, palette = "YlGnBu", rev = TRUE), title = "MFL facilities") + tm_layout(title = "MFL")

tmap_admin2_MFL

# district for each facility, intersect the other way
intersect_admin2_MFL = st_intersection(malawi_facilities_MFL, malawi_admin2)


# WHO, admin1 (dataset already has admin1 column)

# Map with admin1 layer
tmap_admin1_WHO = tmap_facilities_WHO + tm_shape(st_geometry(malawi_admin1)) + tm_borders()
tmap_admin1_WHO

## intersecting for Map
facility_admin1_WHO = st_intersects(malawi_admin1, sf_malawi_WHO, sparse = TRUE)
malawi_admin1$facility_WHO = lengths(facility_admin1_WHO)
head(malawi_admin1)

## Map - number of facilities per region
tmap_admin1_WHO2 = tm_shape(st_geometry(malawi_admin1)) + tm_borders() + tm_shape(malawi_admin1) + tm_fill("facility_WHO", style = "cat", title = "WHO facilities", palette = sequential_hcl(4, palette = "YlGnBu", rev = TRUE))

tmap_admin1_WHO2


# freq of facility types by region
malawi_WHO$Admin1 = as.factor(malawi_WHO$Admin1)
df_facility_admin1_WHO = as.data.frame(table(malawi_WHO$Admin1, 
                                             malawi_WHO$`Facility type`))

# Central
admin1_central = filter(df_facility_admin1_WHO, Var1 == "Central")

admin1_central_plot = ggplot(admin1_central, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "WHO Central Region") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0, 250))

# Southern
admin1_southern = filter(df_facility_admin1_WHO, Var1 == "Southern")

admin1_southern_plot = ggplot(admin1_southern, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "WHO Southern Region") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0,250))

# Northern
admin1_northern = filter(df_facility_admin1_WHO, Var1 == "Northern")

admin1_northern_plot = ggplot(admin1_northern, aes(x=Var2, y=Freq)) + geom_bar(stat = "identity", fill="slategray") + labs(x="Facility type", y="Frequency", title = "WHO Northern Region") + theme(plot.title = element_text(face = "bold")) + theme_minimal() + coord_flip() + expand_limits(y=c(0, 150))


# WHO, admin2 
facility_admin2_WHO = st_intersects(malawi_admin2, sf_malawi_WHO, sparse = TRUE)
malawi_admin2$facility_WHO = lengths(facility_admin2_WHO)

# Map - number of facilities per district 
tmap_admin2_WHO = tm_shape(st_geometry(malawi_admin2)) + tm_borders() + tm_shape(malawi_admin2) + tm_fill("facility_WHO", palette = sequential_hcl(4, palette = "YlGnBu", rev = TRUE), title = "WHO facilities")

tmap_admin2_WHO

# district for each facility, intersect the other way 
intersect_admin2_WHO = st_intersection(sf_malawi_WHO, malawi_admin2)


# final plots
plot_admin1_central = align_plots(admin1_region1_plot, admin1_central_plot, align = "v")
ggdraw(plot_admin1_central[[1]])
ggdraw(plot_admin1_central[[2]])

plot_admin1_southern = align_plots(admin1_region2_plot, admin1_southern_plot, align = "v")
ggdraw(plot_admin1_southern[[1]])
ggdraw(plot_admin1_southern[[2]])

plot_admin1_northern = align_plots(admin1_region3_plot, admin1_northern_plot, align = "v")
ggdraw(plot_admin1_northern[[1]])
ggdraw(plot_admin1_northern[[2]])
```

   
   
   5. Automating the process of sorting facility types into 4 tiers as (Falchetta et al., 2020). Tier 1: dispensary or basic health post; Tier 2: health center; Tier 3: provincial hospital or regional hospital; and Tier 4: central hospital or national hospital.

Ways to apply the Falchetta classification:

- Function that recognises the different facility types and in different languages (downside = similar to manual process, hard to account for all variations)
- Use code/id/reference codes provided with data, do they refer to type of facility? Couldn't find how the MoH codes with the MFL data for Malawi were assigned
- Use capacity data if available? 
   
```{r, message=FALSE, warning=FALSE}
# Falchetta classification with Malawi MFL 
for(i in 1:nrow(malawi_facilities_MFL)) {
   if(malawi_facilities_MFL$TYPE[i] == "Central Hospital"){
      malawi_facilities_MFL[i, "Tier"] = 4
   }
   else if(malawi_facilities_MFL$TYPE[i] == "District Hospital" | malawi_facilities_MFL$TYPE[i] == "Hospital"){
      malawi_facilities_MFL[i, "Tier"] = 3
   }
   else if(malawi_facilities_MFL$TYPE[i] == "Health Centre" | malawi_facilities_MFL$TYPE[i] == "Clinic"){
      malawi_facilities_MFL[i, "Tier"] = 2
   }
   else if(malawi_facilities_MFL$TYPE[i] == "Dispensary" | malawi_facilities_MFL$TYPE[i] == "Health Post"){
      malawi_facilities_MFL[i, "Tier"] = 1
   }
   else{
      malawi_facilities_MFL[i, "Tier"] = 0
   }
}

# With Zambia MFL 
zambia_MFL <- read_csv("https://raw.githubusercontent.com/MOH-Zambia/MFL/master/geography/data/facility_list.csv")
zambia_MFL$facility_type = as.factor(zambia_MFL$facility_type)
levels(zambia_MFL$facility_type)

# https://healthmarketinnovations.org/sites/default/files/Final_%20CHMI%20Zambia%20profile.pdf states 3 broad
# levels,
# Primary = health centres and health posts
# Secondary = Provincial and district hospitals
# Tertiary = Central hospitals
# https://www.severemalaria.org/countries/zambia/zambia-health-system - describes hospital levels

for(i in 1:nrow(zambia_MFL)) {
   if(is.na(zambia_MFL$facility_type[i]) == TRUE){
      zambia_MFL[i, "Tier"] = 0
   }
   else if(zambia_MFL$facility_type[i] == "Hospital - Level 3"){
      zambia_MFL[i, "Tier"] = 4
   }
   else if(zambia_MFL$facility_type[i] == "Hospital - Level 2" | zambia_MFL$facility_type[i] == "Hospital - Level 1"){
      zambia_MFL[i, "Tier"] = 3
   }
   else if(zambia_MFL$facility_type[i] == "Hospital Affiliated Health Centre" | zambia_MFL$facility_type[i] == "Rural Health Centre" |zambia_MFL$facility_type[i] == "Zonal Health Centre" | zambia_MFL$facility_type[i] == "Urban Health Centre"){
      zambia_MFL[i, "Tier"] = 2
   }
   else if(zambia_MFL$facility_type[i] == "Border Health Post" | zambia_MFL$facility_type[i] == "Health Post"){
      zambia_MFL[i, "Tier"] = 1
   }
   else{
      zambia_MFL[i, "Tier"] = 0 
   }
}

```

   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
   
