---
title: "02 Selection"
author: '35762384'
date: "17/02/2021"
output: 
  html_document: 
    keep_md: yes
    code_folding: show  
---  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(sf)
library(readxl)
library(dplyr)
library(plyr)
library(ggplot2)
library(afrihealthsites)
library(ggpubr)
library(afriadmin)
library(tmap)
library(cowplot)
library(forcats)
library(colorspace)
library(readr)
library(waffle)
library(sunburstR)
library(htmltools)
library(htmlwidgets)

# Install Malawi MFL
malawi_MFL = read_excel("~/malawi-health-facilities-1/MHFR_Facilities 1.xlsx")

# Convert to sf

## omit NA's
new_malawi_MFL = na.omit(malawi_MFL)

## transform geometry columns into numeric 
new_malawi_MFL = transform(new_malawi_MFL, LATITUDE = as.numeric(LATITUDE), 
                                           LONGITUDE = as.numeric(LONGITUDE))

## convert to sf object
malawi_facilities_MFL = st_as_sf(new_malawi_MFL, coords = c("LONGITUDE", "LATITUDE"), dim = "XY")

malawi_facilities_MFL = st_set_crs(malawi_facilities_MFL, 4326)

malawi_WHO <- afrihealthsites("malawi", datasource='who', plot=FALSE, returnclass='dataframe')
```
  
##### Figures  

- Columns of data 
```{r}
head(malawi_MFL)
head(malawi_WHO)
```

- Facility plots 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# MFL
# Re-order the facility types
malawi_MFL$TYPE = as.factor(malawi_MFL$TYPE)

malawi_MFL$TYPE = factor(malawi_MFL$TYPE, levels = c("Central Hospital", "District Hospital", "Hospital", "Health Centre", "Clinic", "Health Post", "Dispensary", "Private", "Unclassified"))

# Number of each type + ownership
facility_types_MFL = as.data.frame(table(malawi_MFL$TYPE, malawi_MFL$OWNERSHIP))

# Only showing government, private and CHAM ownership
facility_types_MFL$new_Var2 = revalue(facility_types_MFL$Var2, c("Aquaid Lifeline"="Other", "Non-Government"="Other", "Parastatal"="Other", "Mission/Faith-based (other than CHAM)"="Other"))
facility_types_MFL$new_Var2 = factor(facility_types_MFL$new_Var2, levels = c("Government", "Private", "Christian Health Association of Malawi (CHAM)", "Other"))


## bar plot of no. of facility types 
plot_facility_types_MFL = ggplot(facility_types_MFL, aes(x=Var1, y=Freq, fill=new_Var2)) + geom_bar(position = "stack", stat = "identity")

plot_facility_types_MFL = plot_facility_types_MFL + labs(x = "Facility types", y = "Frequency", fill="Ownership") + scale_fill_brewer(palette = "Set2") + coord_flip() + theme_minimal() + ggtitle("MFL") + theme(axis.title.y = element_blank(), legend.title = element_text(size = 8, face = "bold"), legend.text = element_text(size = 8), plot.title = element_text(face = "bold", hjust = 0), legend.position = "bottom")

# WHO
# Re order facility types and ownership 
malawi_WHO$`Facility type` = as.factor(malawi_WHO$`Facility type`)
malawi_WHO$`Facility type` = factor(malawi_WHO$`Facility type`, levels = c("Central Hospital", "District Hospital", "Mission Hospital", "Rural Hospital", "Community Hospital", "Health Centre", "Clinic", "Health Post/Dispensary"))

malawi_WHO$Ownership = as.factor(malawi_WHO$Ownership)
malawi_WHO$Ownership = factor(malawi_WHO$Ownership, levels = c("MoH", "FBO", "Local authority", "NGO"))

# No. of original facility types + ownership
facility_types_WHO = as.data.frame(table(malawi_WHO$`Facility type`, malawi_WHO$Ownership))

## bar plot of original facility types 
plot_facility_types_WHO = ggplot(facility_types_WHO, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(position = "stack", stat = "identity")
plot_facility_types_WHO = plot_facility_types_WHO + labs(x = "Facility types", y = "Frequency", fill="Ownership") + scale_fill_brewer(palette = "Set2") + coord_flip() + theme_minimal() + ggtitle("WHO") + theme(axis.title.y = element_blank(), legend.title = element_text(size = 8, face = "bold"), legend.text = element_text(size = 8), plot.title = element_text(face = "bold", hjust = 0), legend.position = "bottom") + expand_limits(y=c(0,600))

plot_facilities = align_plots(plot_facility_types_MFL, plot_facility_types_WHO, align = "v")
ggdraw(plot_facilities[[1]])
ggdraw(plot_facilities[[2]])
```

- Maps 
```{r, echo=FALSE, message=FALSE, fig.show='hide', warning=FALSE}
malawi_admin1 = afriadmin("malawi", level = 1, plot = "sf")
malawi_admin2 = afriadmin("malawi",level=2, plot='sf')

new_malawi_WHO = na.omit(malawi_WHO) ## omit NA

sf_malawi_WHO = st_as_sf(new_malawi_WHO, coords = c("Long", "Lat"), dim = "XY")
sf_malawi_WHO = st_set_crs(sf_malawi_WHO, 4326)

tmap_mode("view")
```


```{r, message=FALSE, echo=FALSE, warning=FALSE}
# MFL 
facility_admin2_MFL = st_intersects(malawi_admin2, malawi_facilities_MFL, sparse = TRUE)
malawi_admin2$facility = lengths(facility_admin2_MFL)

tmap_admin2_MFL = tm_shape(st_geometry(malawi_admin2)) + tm_borders() + tm_shape(malawi_admin2) + tm_fill("facility", palette = sequential_hcl(6, palette = "YlGnBu", rev = TRUE), title = "MFL facilities") + tm_layout(title = "MFL")

tmap_admin2_MFL

# WHO
facility_admin2_WHO = st_intersects(malawi_admin2, sf_malawi_WHO, sparse = TRUE)
malawi_admin2$facility_WHO = lengths(facility_admin2_WHO)

tmap_admin2_WHO = tm_shape(st_geometry(malawi_admin2)) + tm_borders() + tm_shape(malawi_admin2) + tm_fill("facility_WHO", palette = sequential_hcl(4, palette = "YlGnBu", rev = TRUE), title = "WHO facilities")

tmap_admin2_WHO

# Point locations - combined 

tmap_admin2 = tm_shape(st_geometry(malawi_admin2)) + tm_borders()

tmap_facilities_WHO = tmap_admin2 + tm_shape(sf_malawi_WHO) + tm_dots(col = "Facility type", palette = "-viridis") + tm_layout(frame = FALSE, asp = 2, title = "WHO", title.position = c("left", "top"))

tmap_facilities_MFL_WHO = tmap_facilities_WHO + tm_shape(malawi_facilities_MFL) + tm_dots(col = "TYPE", palette = "-YlOrRd", alpha = 0.7) + tm_layout(frame = FALSE, asp = 2, title = "MFL & WHO", title.position = c("left", "top"))

tmap_facilities_MFL_WHO
```
  
##### Missing values/wrong coordinates  

- 181 in the MFL (this includes NA's and coordinates that did not intersect with admin levels) and 9 in the WHO data. None are the same facilities/share same names. 
```{r, message=FALSE, warning=FALSE}
# MFL - 1546 before omitting NA's, then 1427.
# 62 coordinates did not intersect with admin data - now 1365

# using anti_join() to see which facilities were omitted 
df_malawi_facilities_MFL = as.data.frame(malawi_facilities_MFL) # convert to data frame to use in anti_join

NA_MFL = anti_join(malawi_MFL, df_malawi_facilities_MFL) # 119 missing coordinates in MFL

# adding the number of wrong coordinates - 61
intersect_admin1_MFL = st_intersection(malawi_facilities_MFL, malawi_admin1) # returns data frame of matches

intersect_admin1_MFL = as.data.frame(intersect_admin1_MFL) # convert to data frame to use in anti_join

no_intersect_admin1_MFL = anti_join(malawi_facilities_MFL, intersect_admin1_MFL) # working out which facilities did not intersect

no_intersect_admin1_MFL_1 = no_intersect_admin1_MFL[ ,-c(10, 11)]
NA_MFL_1 = NA_MFL[, -c(10, 11)]
no_intersect_admin1_MFL_1 = st_set_geometry(no_intersect_admin1_MFL_1, NULL)
names(NA_MFL_1)[3] = "COMMON.NAME"
names(NA_MFL_1)[9] = "DATE.OPENED"

NA_MFL = rbind(NA_MFL_1, no_intersect_admin1_MFL_1) # combined dataframe of NA and wrong coordinates in MFL

# WHO - 648 with 9 missing coordinates, now 639
NA_WHO = anti_join(malawi_WHO, new_malawi_WHO)
```
  
##### Proportion that are privately owned in the MFL data  

- 30.3% of facilities are privately owned (433/1427)  
- Top 3 types are clinics (356), dispensaries (45) and hospitals (16)  
```{r}
private_MFL = filter(malawi_facilities_MFL, OWNERSHIP == "Private")
# 433 private facilities, 30.3%

# which types are private?
private_MFL$TYPE = as.factor(private_MFL$TYPE)
types_private_MFL = as.data.frame(table(private_MFL$TYPE))
types_private_MFL
```

##### Proportion of facility types in each district 
```{r, message=FALSE, warning=FALSE}
# district per facility (does not include the 62 facilities) MFL
intersect_admin2_MFL = st_intersection(malawi_facilities_MFL, malawi_admin2)
intersect_admin2_MFL$TYPE = as.factor(intersect_admin2_MFL$TYPE)
intersect_admin2_MFL$shapeName = as.factor(intersect_admin2_MFL$shapeName)

# types by district
by_district_MFL = as.data.frame(table(intersect_admin2_MFL$TYPE, intersect_admin2_MFL$shapeName))
head(by_district_MFL)

# WHO 
intersect_admin2_WHO = st_intersection(sf_malawi_WHO, malawi_admin2)
intersect_admin2_WHO$Facility.type = as.factor(intersect_admin2_WHO$Facility.type)
intersect_admin2_WHO$shapeName = as.factor(intersect_admin2_WHO$shapeName)

by_district_WHO = as.data.frame(table(intersect_admin2_WHO$Facility.type, intersect_admin2_WHO$shapeName))
head(by_district_WHO)

# Example pie charts for district 

## Balaka 
balaka_MFL = filter(by_district_MFL, Var2 == "Balaka")
balaka_MFL = filter(balaka_MFL, Freq > 0)
balaka_MFL = balaka_MFL[ , -c(2)]

## plot
balaka_plot_MFL = sunburst(balaka_MFL, count = TRUE, legend = list(w=120))
balaka_plot_MFL

## 
balaka_WHO = filter(by_district_WHO, Var2 == "Balaka")
balaka_WHO = filter(balaka_WHO, Freq > 0)
balaka_WHO = balaka_WHO[ , -c(2)]

## plot
balaka_plot_WHO = sund2b(balaka_WHO, rootLabel = "Facilities")
balaka_plot_WHO

## Lilongwe 
lilongwe_MFL = filter(by_district_MFL, Var2 == "Lilongwe")
lilongwe_MFL = filter(lilongwe_MFL, Freq > 0)
lilongwe_MFL = lilongwe_MFL[ , -c(2)]

## plot
lilongwe_plot_MFL = sund2b(lilongwe_MFL, rootLabel = "Facilities")
lilongwe_plot_MFL

## 
lilongwe_WHO = filter(by_district_WHO, Var2 == "Lilongwe")
lilongwe_WHO = filter(lilongwe_WHO, Freq > 0)
lilongwe_WHO = lilongwe_WHO[ , -c(2)]

## plot
lilongwe_plot_WHO = sund2b(lilongwe_WHO, rootLabel = "Facilities")
lilongwe_plot_WHO
```
  
##### Duplicates within dataset  

- 22 names in MFL appear more than once, of these Mpepa Health Centre (2 decimal places), Liwonde Medical Clinic (identical), Ngatala Health Post (3 decimal places), Cape Maclear (identical) and LIFE STYLE PVT CLINIC (identical) have similar/same coordinates 
- Of these, Ngatala, Cape Maclear and LIFE STYLE PVT CLINIC are most likely the same entries as other attributes are also the same. The other facilities differ in type and so might not be duplicates. 
- 10 names in WHO appear more than once, Mkango Health Centre has same coordinates (4 decimal places) and other attributes are also the same  

- Duplicate coordinates: in the MFL, 44 coordinates have been repeated at least more than once.  
Interesting note is that 24 facilities in Blantyre have reused the same coordinates (34.3015278, -13.2512161) and they consist of 1 hospital and a mix of clinics and health centres.  All have unique names as well.  
Same coordinates (33.7895208991213, -14.0017665100358) have also been used for 4 facilities, all in different districts.  
Some results are due to no coordinates being available and (-1,1) for example is inputted instead.  
No duplicates in the WHO data  

- Duplicates between datasets?  
```{r}
# MFL
duplicate_name_MFL = malawi_facilities_MFL[malawi_facilities_MFL$NAME %in% malawi_facilities_MFL$NAME[duplicated(malawi_facilities_MFL$NAME)], ]

# WHO 
duplicate_name_WHO = new_malawi_WHO[new_malawi_WHO$`Facility name` %in% new_malawi_WHO$`Facility name`[duplicated(new_malawi_WHO$`Facility name`)], ]

# Same coordinates within dataset

## MFL 
duplicate_coord_MFL = malawi_facilities_MFL[malawi_facilities_MFL$geometry %in% malawi_facilities_MFL$geometry[duplicated(malawi_facilities_MFL$geometry)], ]

## WHO 
sf_malawi_WHO$geometry[duplicated(sf_malawi_WHO$geometry)] # none 

# Blantyre 
blantyre_duplicate_coord_MFL = filter(duplicate_coord_MFL, DISTRICT == "Blantyre")
blantyre_MFL = filter(by_district_MFL, Var2 == "Blantyre")
blantyre_MFL = filter(blantyre_MFL, Freq > 0)
blantyre_MFL = blantyre_MFL[ , -c(2)]

## plot
blantyre_plot_MFL = sund2b(blantyre_MFL, rootLabel = "Facilities")
blantyre_plot_MFL

```

##### Facilities within 50km
```{r}

```
















